### 2023-03-28 definitions.tex
details

Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo. In such \emph{Finite Antichains} sets $(X,\leq)$, every downward closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_i$ (an ideal is a directed downward closed subset): $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.

In what follows, a downward closed set is represented by its finite set of ideals 
%		$D=\downarrow D$   $I_D=\{J_1, J_2,...,J_n\}$ 		satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ 
and an upward closed set $U$ is represented by its finite set minimal elements.
% $M_U$ of its  satisfying $U=\uparrow M_U$.

@@

\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
A {\em (resp. strongly) well-structured transition system} (abbreviated as WSTS)  $(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ between states such that  
%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
the transition relation $ \rightarrow$ is compatible (resp. strongly compatible) with the wqo, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$ , there exists 
	$t_2 \in S$ with 
	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$ 			
				(resp. $s_2 \rightarrow t_2$ ).
\end{definition}

@@

\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ and $\pred^*(U )$ are upward-closed.
\end{proposition}
%\alain{étrange ces deux propositions, l'une suffirait}
\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.

@@

\begin{claim}{(stability of upward-closed sets)}
Let $U, V \subseteq S$ be upward-closed. Then the sets $U \cup V$, and $U \cap V$ are upward-closed.
\end{claim}

@@

\begin{fact}\label{fact basis}
% Fact 3 ([1]). 
% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
For every upward-closed set $U \subseteq S$, there exists a finite basis $B$ of $U$. 
% (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
\end{fact}

\begin{definition}{ (index)}. 
If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is upward-closed and $k \geq 0$, let $U^k= \bigcup_{0 \leq j \leq k} \pred^j(U)$.
The {\em index} $k(U)$ is the
smallest $k_0$ s.t. $U^k = U^{k_0}$ for all $k \geq k_0$.
\end{definition}

If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set, the sets $\pred(U )$, and $\pred^{\geq k}(U )$ for
every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(U)$.

\begin{remark}
$U^{k+1} $ can be rewritten $U^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(U) = 
U \cup \bigcup_{1 \leq j \leq k+1} \pred^j(U) =
U \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(U))
=  U \cup \pred(U^k)$.
\end{remark}

@@

Since a finite basis of $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $U^{k+1}$ . By
Fact~\ref{fact basis}, we can compute a basis of $U^{k+1}$.
\end{proof}

\begin{theorem}
A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward closed set $U \subseteq S$ given with its finite basis $B_U$.
\end{theorem}

>>>>

Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo. In such \emph{Finite Antichains} sets $(X,\leq)$, every downward closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ (an ideal is a directed downward closed subset) such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.

In what follows, a downward closed set is represented by its finite set of ideals 
%		$D=\downarrow D$   $I_D=\{J_1, J_2,...,J_n\}$ 		satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ 
and an upward closed set $U$ is represented by its finite set of minimal elements.
% $M_U$ of its  satisfying $U=\uparrow M_U$.

@@

\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
A {\em well-structured transition system} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ such that  
%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
the transition relation $ \rightarrow$ is compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

We say that a WSTS $\mathscr{S}$ has the \emph{strong compatibility} when moreover for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_1 \leq t_2$ and $s_2 \rightarrow t_2$.

@@

\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed.
\end{proposition}
%\alain{étrange ces deux propositions, l'une suffirait}
\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.

@@

%
\iffalse
%
\begin{claim}{(stability of upward-closed sets)}
Let $U, V \subseteq S$ be upward-closed. Then the sets $U \cup V$, and $U \cap V$ are upward-closed.
\end{claim}
%
\fi
%

@@

%			\begin{fact}\label{fact basis}
% Fact 3 ([1]). 
% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,

%For every upward-closed set $U \subseteq S$, there exists a finite basis $B$ of $U$.

% (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
%			\end{fact}

For a WSTS $\mathscr{S}$ and an upward closed set $U$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U \cup \pred(U_{k-1})$ for $k \geq 1$. When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)$ converges. Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$. We then have:  $\pred^*(U) = U_{k_0}$.

\alain{ me semble maintenant inutile
%
\begin{definition}{ (index)}. 
If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is upward-closed and $k \geq 0$, let $U_k= \bigcup_{0 \leq n \leq k} \pred^n(U)$.
The {\em index} $k(U)$ is the
smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$.
\end{definition}
%
%
If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set, the sets $\pred(U )$, and $\pred^{\geq k}(U )$ for
every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(U)$.
%
\begin{remark}
$U^{k+1} $ can be rewritten $U^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(U) = 
U \cup \bigcup_{1 \leq j \leq k+1} \pred^j(U) =
U \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(U))
=  U \cup \pred(U^k)$.
\end{remark}

@@

Since a finite basis of $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $U^{k+1}$ . By
Fact~\ref{fact basis}, we can compute a basis of $U^{k+1}$.
\end{proof}}

\begin{theorem}(ref)
A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward closed set $U \subseteq S$ given with its finite basis $pb(U)$.
\end{theorem}
